Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-14 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, we can define
a QIVPM mapping to $\widehat{\mathscr{I}}_{n}$ based on any Born
rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\widehat{\mathscr{I}}_{n}$ are overlapping, if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)$ are contained in both intervals~$[\ell_{1},r_{1}]$
and $[\ell_{2},r_{2}]$, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
will be $[\ell_{i},r_{i}]$ if $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$. When there is a tie,
we set $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
for even $k$. For example, when $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{2} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
It is straightforward to verify $\widehat{\iota}_{2}\circ\mu_{1}$
is a QIVPM. It will be denoted by $\bar{\mu}_{4}$ whose values on
the minimal subspace of events containing $\proj{0}$, $\proj{1}$,
and $\proj{2}$, $\events_{C}$, is computed by Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\bar{\mu}_{4}$ based on $\mu_{1}$
on $\events_{C}$.}
\begin{tabular}{ccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\bar{\mu}_{4}\left(P\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}

The QIVPM~$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is well-behavior
in a sense that the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need some more technical
notions. We first define a notion of distance based on the following
norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
Then, for any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
any other state~$\rho'$ consistent with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
must close to $\rho$ in a sense that $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$.
Reversely, if a state~$\rho'$ is close to $\rho$ in a sense that
$\left\Vert \rho-\rho'\right\Vert \le1/n$, then $\rho'$ is consistent
with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.

>>>>

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by the interval containing them if we carefully pick a
set of overlapping intervals. For example, consider the following
collection of intervals: 
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Then, we can define
a QIVPM mapping to $\widehat{\mathscr{I}}_{n}$ based on any Born
rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\widehat{\mathscr{I}}_{n}$ are overlapping, if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)$ are contained in both intervals~$[\ell_{1},r_{1}]$
and $[\ell_{2},r_{2}]$, $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
will be $[\ell_{i},r_{i}]$ if $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$. When there is a tie,
we set $\widehat{\iota}_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
for even $k$. For example, when $n=2$, 
\begin{eqnarray}
\widehat{\mathscr{I}}_{2} & = & \set{\left[\frac{k-1}{4},\frac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\frac{1}{4}\right],\left[0,\frac{1}{2}\right],\left[\frac{1}{4},\frac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\frac{1}{2},1\right],\left[\frac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
It is straightforward to verify $\widehat{\iota}_{2}\circ\mu_{1}$
is a QIVPM whose values on the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, is computed
by Table~\ref{tab:Compute-the-QIVPM}. 
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\widehat{\iota}_{2}\circ\mu_{1}$
based on $\mu_{1}$ on $\events_{C}$.}
\begin{tabular}{ccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\widehat{\iota}_{2}\left(\mu_{1}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}

Given any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
we know not only $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, but also the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need to define a notion
of distance based on the following norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
If a state~$\rho'$ is consistent with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $1/\left(12n\right)$. Reversely, if the distance between
$\rho$ and $\rho'$ is bounded by $1/n$, then $\rho'$ is consistent
with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.
